---
title: 'Twist 2018: Airtraffic challenge'
output:
 prettydoc::html_pretty:
   theme: cayman
   highlight: github
   toc: true
   number_sections: true
   df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, fig.align = "center")
```

# Libraries and data
```{r}
rm(list = ls())
library(tidyverse)
library(lubridate)
library(gridExtra)
library(ggfortify)
```

```{r}
# flightdata <- readRDS("twist_zrh/twist_zrh.RDS")
# 
# fac2num <- function(x) as.numeric(as.character(x))
# flightdata <- flightdata %>%
#   mutate(date = as.Date(flightdata$date, "%d.%m.%Y")) %>%
#   mutate_at(vars(lightnings_hour_n, lightnings_hour_f,
#                  winddir_h, windspeed_avg_h, windspeed_peak_h,
#                  global_rad_avg_h, global_rad_avg_h, airpres,
#                  sunshine_dur_min, temp_avg, temp_min,
#                  temp_max, rel_humid, precip), fac2num)
# glimpse(flightdata)
# write_rds(flightdata, "twist_zrh_cleaned.RDS")
flightdata <- readRDS("twist_zrh_cleaned.RDS")
flightdata <- na.omit(flightdata)

flightdata <- flightdata %>%
  mutate(month = month(date),
         hour = hour(planed_time),
         continent = as.character(continent)) %>%
  mutate(month = as.factor(month),
         hour = as.factor(hour),
         continent = as.factor(continent))
glimpse(flightdata)

flightdata <- flightdata %>%
  mutate(delayed = ifelse(abs(as.numeric(diff_in_secs)) > 1800, 1, 0))
flightdata_landing <- flightdata %>%
  filter(start_landing == "L")
flightdata_starting <- flightdata %>%
  filter(start_landing == "S")
```

# EDA visualization
## Looking at time difference distribution
```{r}
plot_all <- flightdata %>%
  ggplot(aes(x = diff_in_secs)) + 
    geom_histogram(bins = 60, col = 1) +
    labs(title = "starting flights: full dataset") +
    facet_wrap(~ start_landing)

plot_withoutoutliers <- flightdata %>%
  filter(abs(diff_in_secs) < 10000) %>%
  ggplot(aes(x = diff_in_secs)) + 
    geom_histogram(bins = 60, col = 1) +
    labs(title = "absolute difference < 10000") +
    facet_wrap(~ start_landing)

grid.arrange(plot_all, plot_withoutoutliers, ncol = 2)
```

## Boxplots for different months and hours of the day
```{r}
flightdata %>%
  filter(abs(diff_in_secs) < 5000) %>%
  ggplot(aes(x = month, y = as.numeric(diff_in_secs))) + 
    geom_boxplot() +
    facet_wrap(~ start_landing)
flightdata %>%
  filter(abs(diff_in_secs) < 5000) %>%
  ggplot(aes(x = hour, y = as.numeric(diff_in_secs))) + 
    geom_boxplot() +
    facet_wrap(~ start_landing)
flightdata_landing <- flightdata %>%
  filter(start_landing == "L")
flightdata_starting <- flightdata %>%
  filter(start_landing == "S")
```

## Comparing Schengen to non-Schengen
```{r}
flightdata %>%
  ggplot(aes(x = schengen, y = as.numeric(diff_in_secs))) +
    geom_boxplot() +
    facet_wrap(~ start_landing) +
    ylim(-5000, 5000) +
    labs(y = "Difference in seconds")
```


## Airlines and airplane types
```{r}
flightdata %>%
  arrange(airline_name) %>%
  group_by(airline_name) %>%
  filter(n() > 1000) %>%
  ggplot(aes(x = airline_name, y = as.numeric(diff_in_secs), 
                         group = airline_name)) +
    geom_boxplot() +
    geom_hline(yintercept = 0, col = 4) +
    facet_wrap(~ start_landing) + 
    ylim(-5000, 10000) +
    coord_flip() +
    labs(x = "Airline name", y = "Difference in seconds", 
         title = "Airlines with more than 1000 flights")

flightdata %>%
  arrange(airplane_type) %>%
  group_by(airplane_type) %>%
  filter(n() > 100) %>%
  ggplot(aes(x = airplane_type, y = as.numeric(diff_in_secs), 
                         group = airplane_type)) +
    geom_boxplot() +
    geom_hline(yintercept = 0, col = 4) +
    facet_wrap(~ start_landing) +
    ylim(-5000, 10000) +
    coord_flip() +
    labs(x = "Airplane type", y = "Difference in seconds", 
         title = "Airplane types with more than 100 flights")
```

# Modeling
## Linear model with quantiative covariates  
```{r}
model1 <- as.formula(as.numeric(diff_in_secs) ~ distance_km + windspeed_avg_h
                     + temp_avg + rel_humid + lightnings_hour_n + lightnings_hour_f
                     + sunshine_dur_min + airpres + global_rad_avg_h)
fit1 <- lm(model1, flightdata_starting)
summary(fit1)
autoplot(fit1)
```

Residuals clearly not Gaussian!

## Best subset model selection with BIC criterion  
```{r}
library(leaps)
modelall <- as.formula(as.numeric(diff_in_secs) ~ distance_km + windspeed_avg_h
                     + temp_avg + rel_humid + lightnings_hour_n + lightnings_hour_f
                     + sunshine_dur_min + airpres + global_rad_avg_h + temp_max
                     + temp_min + precip + windspeed_peak_h + windspeed_peak_h)
modelselection <- regsubsets(modelall, data = flightdata_starting, nvmax = 14)
minmodel <- which.min(summary(modelselection)$bic)
bestmodel <- as.formula(paste(c("as.numeric(diff_in_secs) ~ ", 
                                paste(names(coef(modelselection, minmodel)[-1]), 
                                      collapse = "+")), collapse = ""))
bestfit <- lm(bestmodel, flightdata_starting)
summary(bestfit)
autoplot(bestfit)
```

8 out of 9 variables are included, but the residuals are still not looking Gaussian!


## Logistic regression model using delay > 30 min as outcome
```{r}
model_glm <- as.formula(delayed ~ distance_km + windspeed_avg_h + schengen
                     + temp_avg + rel_humid + lightnings_hour_n + lightnings_hour_f
                     + sunshine_dur_min + airpres + global_rad_avg_h + temp_max
                     + temp_min + precip + windspeed_peak_h + windspeed_peak_h)
fit_glm1 <- glm(model_glm, data = flightdata_starting, family = "binomial")
summary(fit_glm1)

# missclassification
mean(flightdata_starting$delayed != as.numeric(predict(fit_glm1, type = "response") > 0.5))
```

## Adding months and hours to models
```{r}
bestmodel_withtime <- update(bestmodel, . ~ . + month + hour)
fit_time <- lm(bestmodel_withtime, flightdata_starting)
summary(fit_time)
autoplot(fit_time)
```

Ok, linear regression performance is bad and the residuals are clearly not Gaussian, we should use logistic regression from now on!

```{r}
glm_month <- as.formula(delayed ~ distance_km + windspeed_avg_h + schengen
                     + temp_avg + rel_humid + lightnings_hour_n + lightnings_hour_f
                     + sunshine_dur_min + airpres + global_rad_avg_h + temp_max
                     + temp_min + precip + windspeed_peak_h + month)
fit_glm_month <- glm(glm_month, data = flightdata_starting, family = "binomial")
summary(fit_glm_month)
mean(flightdata_starting$delayed != as.numeric(predict(fit_glm_month, type = "response") > 0.5))

glm_month_hour <- as.formula(delayed ~ distance_km + windspeed_avg_h + schengen
                     + temp_avg + rel_humid + lightnings_hour_n + lightnings_hour_f
                     + sunshine_dur_min + airpres + global_rad_avg_h + temp_max
                     + temp_min + precip + windspeed_peak_h + month + hour)
fit_glm_month_hour <- glm(glm_month_hour, data = flightdata_starting, family = "binomial")
summary(fit_glm_month)
mean(flightdata_starting$delayed != as.numeric(predict(fit_glm_month_hour, type = "response") > 0.5))

anova(fit_glm_month, fit_glm_month_hour, test = "LR")
```

## Getting cross-validation missclassification rate
```{r}
set.seed(123)
Kfold_cv_glm <- function(model_form, data, K){
  n <- nrow(data)
  ind_permuted <- sample(n, n, replace = FALSE)
  folds <- cut(ind_permuted, K, labels = FALSE)
  missclass <- sapply(seq(2), function(k){
    train <- which(folds != k)
    fit <- glm(model_form, data = data[train,], family = "binomial")
    mean(data$delayed[-train] != as.numeric(predict(fit, newdata = data[-train,], type = "response") > 0.5))
  })
  return(missclass)
}
missclassification_cv <- Kfold_cv_glm(glm_month_hour, flightdata_starting, 10)
mean(missclassification_cv)
```

## Plotting model predictions
```{r fig.height = 5}
library(visreg)
distance <- visreg(fit_glm_month_hour, "distance_km", scale = "response", gg = TRUE) + ylim(0, 0.6)
windspeed <- visreg(fit_glm_month_hour, "windspeed_avg_h", scale = "response", gg = TRUE) + ylim(0, 0.6)
airpressure <- visreg(fit_glm_month_hour, "airpres", scale = "response", gg = TRUE) + ylim(0, 0.6)
lightning <- visreg(fit_glm_month_hour, "lightnings_hour_f", scale = "response", gg = TRUE) + ylim(0, 0.6)
grid.arrange(distance, windspeed, airpressure, lightning, ncol = 2)

sunshine <- visreg(fit_glm_month_hour, "sunshine_dur_min", scale = "response", gg = TRUE) + ylim(0, 0.6)
windspeed <- visreg(fit_glm_month_hour, "windspeed_avg_h", scale = "response", gg = TRUE) + ylim(0, 0.6)
global_radiation <- visreg(fit_glm_month_hour, "global_rad_avg_h", scale = "response", gg = TRUE) + ylim(0, 0.6)
months <- visreg(fit_glm_month_hour, "month", scale = "response", gg = TRUE) + ylim(0, 0.6)
grid.arrange(sunshine, windspeed, global_radiation, months, ncol = 2)

hours <- visreg(fit_glm_month_hour, "hour", scale = "response", gg = TRUE) + ylim(0, 0.6)
precip <- visreg(fit_glm_month_hour, "precip", scale = "response", gg = TRUE) + ylim(0, 0.6)
lightnings_hour_n <- visreg(fit_glm_month_hour, "lightnings_hour_n", scale = "response", gg = TRUE) + ylim(0, 0.6)
rel_humid <- visreg(fit_glm_month_hour, "rel_humid", scale = "response", gg = TRUE) + ylim(0, 0.6)
grid.arrange(hours, precip, lightnings_hour_n, rel_humid, ncol = 2)
```




